---
title: "GLAM exam report Kieran Schubert"
output: 
  pdf_document:
    number_sections: true
fontsize: 10pt

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/kieranschubert/Desktop/Statistics/2nd_Semester/GLAM/Exam")
#install.packages("statmod", repos='http://cran.us.r-project.org')
library(statmod)
#install.packages("brglm", repos='http://cran.us.r-project.org')
library(brglm)
#install.packages("robustbase", repos='http://cran.us.r-project.org')
library(robustbase)
#install.packages("knitr", repos='http://cran.us.r-project.org')
library(knitr)
#install.packages("gridExtra", repos='http://cran.us.r-project.org')
library(gridExtra)
#install.packages("R.utils", repos='http://cran.us.r-project.org')
library(R.utils)
library(mgcv)
```

# Exploratory Data Analysis

## Dataset Structure

### Dataset size
```{r, echo=F, fig.cap=T}
orange <- read.table("orange_juice.dat", header=T)
y <- orange$CRA8132

data.frame(
  nrows = nrow(orange),
  ncols = ncol(orange)
)
```

### Variable Type
```{r, echo=F}
str(orange) 
```

### Variable range
```{r, echo=F}
data.frame(
  CRA8132.range = range(orange$CRA8132),
  pH.range = range(orange$pH),
  LL.range = range(orange$LL),
  Temp.range = range(orange$Temp),
  Brix.range = range(orange$Brix),
  row.names=c("min", "max")
)
```
### Correlation & Variable Frequency
```{r, echo=F}
cor(orange)
```

```{r, echo=F}
kable(list(table(orange$CRA8132, dnn="CRA8132"), table(orange$pH, dnn="pH"), table(orange$LL, dnn="LL"), table(orange$Temp, dnn="Temp"), table(orange$Brix, dnn="Brix")))
```

```{r, echo=F, eval=F}
par(mfrow=c(2,3))
hist(orange$CRA8132, xlab="Response", main=NULL)
hist(orange$pH, xlab="pH", main=NULL)
hist(orange$LL, xlab="LL", main=NULL)
hist(orange$Temp, xlab="Temp", main=NULL)
hist(orange$Brix, xlab="Brix", main=NULL)
```


```{r, echo=F, eval=F}
# Response vs covariates
par(mfrow=c(2,2))
plot(y, orange$pH, xlab="Response", ylab="pH")
plot(y, orange$LL, xlab="Response", ylab="LL")
plot(y, orange$Temp, xlab="Response", ylab="Temp")
plot(y, orange$Brix, xlab="Response", ylab="Brix")
```

```{r, echo=F,eval=F}
# covariates vs covariates
par(mfrow=c(2,3))
plot(orange$pH, orange$LL, xlab="pH", ylab="LL")
plot(orange$pH, orange$Temp, xlab="pH", ylab="Temp")
plot(orange$pH, orange$Brix, xlab="pH", ylab="Brix")
plot(orange$LL, orange$Temp, xlab="LL", ylab="Temp")
plot(orange$LL, orange$Brix, xlab="LL", ylab="Brix")
plot(orange$Temp, orange$Brix, xlab="Temp", ylab="Brix")
```
\newpage
# Modelling

## Model 1: Full model
```{r, echo=F}
mod1 <- glm(CRA8132~., data=orange, family=binomial)
sum1 <- summary(mod1) # remove params in Deviance for df
sum1
```

```{r, echo=F, include=F}
# GOODNESS OF FIT TEST
mod1.p <- pchisq(sum1$deviance, df=sum1$df.residual, lower.tail=FALSE) #DO NOT REJECT H0 => B_hat != 0 => model significant
```

## Model 2: Removing obs 7 and 44
```{r, echo=F}
orange.omit <- orange[-c(7, 44),]
mod2  <- glm(CRA8132~.,family=binomial,data=orange.omit)
sum2 <- summary(mod2)
sum2
# all coefs significant except Brix
# AIC is much lower than full model (39.82)
```

```{r, echo=F, include=F}
mod2.p <- pchisq(sum2$deviance, df=sum2$df.residual, lower.tail=FALSE) 
```

## Model 3 : Removing Brix
```{r, echo=F}
mod3 <-glm(CRA8132~pH+LL+Temp,family=binomial,data=orange.omit)
sum3 <- summary(mod3)
sum3
```

```{r, echo=F, include=F}
mod3.p <- pchisq(sum3$deviance, df=sum3$df.residual, lower.tail=FALSE) 
```

## Model 4: GAM
```{r, fig.height=2.5, fig.width=7, echo=F}
par(mfrow = c(2, 2), mar=c(1, 2, 1, 1) + 0.1, xpd=NA)
mod2.gam <- gam(CRA8132~s(pH, k=4)+s(LL, k=4)+s(Temp, k=4)+s(Brix, k=4), family=binomial, data=orange)
plot(mod2.gam,cex=3,residuals=T,pages=1)

mod4 <- glm(CRA8132~pH+LL+Temp+I(Brix^2), family=binomial, data=orange)
sum4 <- summary(mod4)
#sum7
```

```{r, echo=F, include=F}
mod4.p <- pchisq(sum4$deviance, df=sum4$df.residual, lower.tail=FALSE) 
```
## Model 5: Exhaustive Search (with all observations, 1st order interactions, Brix^2)

```{r, echo=F, warning=FALSE}
# Exhaustive search, all logistic regression models with interactions, LOGIT link
pH_LL <- orange$pH*orange$LL
pH_Temp <- orange$pH*orange$Temp
pH_Brix <- orange$pH*orange$Brix
LL_Temp <- orange$LL*orange$Temp
LL_Brix <- orange$LL*orange$Brix
Temp_Brix <- orange$Temp*orange$Brix
Brix2 <- orange$Brix^2

orange.aug <- cbind(orange, pH_LL, pH_Temp, pH_Brix, LL_Temp, LL_Brix, Temp_Brix, Brix2)
regressors <- colnames(orange.aug)[-5]
regMat <- expand.grid(c(TRUE,FALSE), c(TRUE,FALSE),
                      c(TRUE,FALSE), c(TRUE,FALSE), 
                      c(TRUE,FALSE), c(TRUE,FALSE),
                      c(TRUE,FALSE), c(TRUE,FALSE),
                      c(TRUE,FALSE), c(TRUE,FALSE),
                      c(TRUE,FALSE))

regMat <- regMat[-(dim(regMat)[1]),] #The last row is the null model so we remove it
names(regMat) <- regressors

allModelsList <- apply(regMat, 1, function(x) as.formula(paste(c("CRA8132~1", regressors[x]), collapse="+")))

allModelsResults <- lapply(allModelsList, function(x) glm(x, data=orange, family=binomial))

library(plyr)
aic <- ldply(allModelsResults, function(x) as.data.frame(
                     t(AIC(x))))
#plot(aic$V1, type="o")
#sort(aic$V1, index.return=T)

allModelsResults[1201]
mod5 <- glm(CRA8132~pH+LL+Temp+Brix+pH_Brix+LL_Brix+Temp_Brix, data=orange.aug, family=binomial)
sum5 <- summary(mod5)
sum5
```

```{r, echo=F, include=F}
step(mod5)

#full model has lowest AIC (62.33) => include all predictors 
```

```{r, echo=F, include=F}
mod5.p <- pchisq(sum5$deviance, df=sum5$df.residual, lower.tail=FALSE) 
```

## Model 6: bias reduction glm without obs 41
```{r, echo=F, fig.height=2.5, fig.width=7, warning=F}
data <- orange.aug[-c(41),]
mod <- glm(CRA8132~pH+LL+Temp+Brix+pH_Brix+LL_Brix+Temp_Brix, data=data, family=binomial)
sum <- summary(mod)
#sum

mod6 <- brglm(CRA8132~pH+LL+Temp+Brix+pH_Brix+LL_Brix+Temp_Brix, data=data, family=binomial)
sum6 <- summary(mod6)
sum6

```

```{r, echo=F, include=F}
mod6.p <- pchisq(sum6$deviance, df=sum6$df.residual, lower.tail=FALSE) 
```

## Model comparison

### Model Formulas
```{r, echo=F}
paste("Model 1: ", formula(mod1))[3]
paste("Model 2: ", formula(mod2))[3]
paste("Model 3: ", formula(mod3))[3]
paste("Model 4: ", formula(mod4))[3]
paste("Model 5: ", formula(mod5))[3]
paste("Model 6: ", formula(mod6))[3]

```

### Coefficients
```{r, echo=F}

# Model 1: full model, all observations
# Model 2: model without obs. 7 and 44
# Model 3: model without Brix, without obs. 7 and 44
# Model 4: exhaustive search model, with interactions, all obs
# Model 5: exhaustie search model, with interactions, without obs. 7 and 44
# Model 6: bias reduction glm
# model 7: Quadratic fit for Brix

#Coefficients
data.frame(
  predictors.mod1 = names(coef(mod1)),
  beta.mod1 = c(coef(mod1)), 
  predictors.mod2 = names(coef(mod2)),
  beta.mod2 = c(coef(mod2)),
  predictors.mod3 = c(names(coef(mod3)), NA),
  beta.mod3 = c(coef(mod3), NA),
  predictors.mod4 = names(c(coef(mod4))),
  beta.mod4 = c(coef(mod4)),
  row.names=NULL
  )
data.frame(
  predictors.mod5 = names(coef(mod5)),
  beta.mod5 = c(coef(mod5)),
  predictors.mod6 = names(coef(mod6)),
  beta.mod6 = c(coef(mod6))
)

```

### Standard Errors
```{r, echo=F}
#Std errors
data.frame(
  predictors.mod1 = names(coef(mod1)),
  stderr.mod1 = c(summary(mod1)$coef[,2]), 
  predictors.mod2 = names(coef(mod2)),
  stderr.mod2 = c(summary(mod2)$coef[,2]),
  predictors.mod3 = c(names(coef(mod3)), NA),
  stderr.mod3  = c(summary(mod3)$coef[,2], NA),
  predictors.mod4 = names(coef(mod4)),
  stderr.mod4 = c(summary(mod4)$coef[,2]),
  row.names=NULL
)

data.frame(
    predictors.mod5 = names(coef(mod5)),
    stderr.mod5 = c(summary(mod5)$coef[,2]),
    predictors.mod6 = names(coef(mod6)),
    stderr.mod6 = c(summary(mod6)$coef[,2])
)
```

### AIC
```{r, echo=F}
#AIC => use because based on log-lik not deviance

data.frame(
  mod1.aic = summary(mod1)$aic,
  mod2.aic = summary(mod2)$aic,
  mod3.aic = summary(mod3)$aic,
  mod4.aic = summary(mod4)$aic,
  mod5.aic = summary(mod5)$aic,
  mod6.aic = summary(mod6)$aic
)


#par(mfrow =c(2,2))
#plot(lr.full.logit.fit, which=c(1,2), main = "lr1 logit")
#plot(brglm.fit, which=c(1,2), main = "brglm logit")


#anova(lr.full.logit.fit, lr4.probit.fit, test="Chisq")
#anova(lr.full.logit.fit, test="Chisq") #sequential analysis of deviance test => compare intercept model with models with added variables (nested) => DEPENDS ON ORDER OF THE VARIABLES, IN GLM FORMULA

#compare models with deviance (LRT): NESTED MODELS p.114 => stepwise selection, start from full or null model
#pchisq(cps.fit1$null.dev-cps.fit1$dev,
#       df=cps.fit1$df.null-cps.fit1$df.res,
#       lower.tail=F)
```

### Model p-values
```{r, echo=F}
data.frame(
  mod1 = mod1.p,
  mod2 = mod2.p,
  mod3 = mod3.p,
  mod4 = mod4.p,
  mod5 = mod5.p,
  mod6 = mod6.p
)
```

## Residuals and Model Checking of Final Model (Model 5)
```{r, echo=F, fig.height=2.5, fig.width=7}
final.mod <- mod5
# RESIDUAL ANALYSIS p.98!!!!!!!
par(mfrow = c(2, 2), mar=c(1, 2, 1, 1) + 0.1)
plot(final.mod) # obs 7, 44 influential
```

### Pearson/Deviance residuals
```{r, echo=F, fig.height=2.5, fig.width=7}
par(mfrow = c(2, 2), mar=c(1, 2, 1, 1) + 0.1, xpd=NA)
plot(predict(final.mod), residuals(final.mod, type="pearson"), ylab="pearson resids", xlab="fitted values")
plot(predict(final.mod), residuals(final.mod, type="deviance"), ylab="deviance resids", xlab="fitted values")
plot(sort(predict(final.mod)), sort(residuals(final.mod, type="pearson")), ylab="pearson resids", xlab="fitted values")
plot(sort(predict(final.mod)), sort(residuals(final.mod, type="deviance")), ylab="deviance resids", xlab="fitted values")
```

### Randomized quantile residuals
```{r, echo=F, fig.height=3, fig.width=6}
#Randomized quantile residuals
par(mfrow=c(1,1))
rqresid.final.mod <- qresid(final.mod)
qqnorm(rqresid.final.mod)
qqline(rqresid.final.mod)
```

### Residuals vs covariates
```{r, echo=F, fig.height=4, fig.width=10}
#Residuals vs covariates
final.mod.res <- residuals(final.mod,type='deviance')

par(mfrow = c(2, 2), mar=c(1, 2, 1, 1) + 0.1, xpd=NA)
plot(jitter(orange$pH), final.mod.res, main = "pH")
abline(h=0, lty = 2)
lines(smooth.spline(orange$pH, final.mod.res),col='limegreen',lwd=2)

plot(jitter(orange$LL), final.mod.res, main = "LL") 
lines(smooth.spline(orange$LL, final.mod.res,df=4),col='limegreen',lwd=2)
abline(h=0, lty = 2)

plot(jitter(orange$Temp), final.mod.res, main = "Temp") 
lines(smooth.spline(orange$Temp, final.mod.res,df=4),col='limegreen',lwd=2)
abline(h=0, lty = 2)

plot(jitter(orange$Brix), final.mod.res, main = "Brix") 
lines(smooth.spline(orange$Brix, final.mod.res,df=4),col='limegreen',lwd=2)
abline(h=0, lty = 2)
```

## Interpretation

```{r, echo=F, eval=T, fig.height=3, fig.width=5}
pH.grid = round(seq(min(orange$pH), max(orange$pH), length.out=5), 2)

LL.grid = round(seq(min(orange$LL), max(orange$LL), length.out=5), 2)

Brix.grid = round(seq(min(orange$Brix), max(orange$Brix), length.out=5), 2)

Temp.grid = round(seq(min(orange$Temp), max(orange$Temp), length.out=5), 2)

#TEMP FIXED
# PROBABILITY TABLE: Temp=25
grid <- expand.grid(pH=pH.grid, LL=LL.grid, Temp=25)
table1 <- round(matrix(predict(mod3, newdata=grid, type="response"), nrow=5, ncol=5), 2)
colnames(table1) <- paste("pH", pH.grid, collape=" ")
rownames(table1) <- paste("LL", LL.grid, collape=" ")
#kable(table, caption="temp=25")

# PROBABILITY TABLE: Temp=35
grid <- expand.grid(pH=pH.grid, LL=LL.grid, Temp=43)
table2 <- round(matrix(predict(mod3, newdata=grid, type="response"), nrow=5, ncol=5), 2)
colnames(table2) <- paste("pH", pH.grid, collape=" ")
rownames(table2) <- paste("LL", LL.grid, collape=" ")
#kable(table, caption="temp=35")
kable(cbind(table1, table2), caption="Temp=25 and Temp=35")

# PROBABILITY TABLE: Temp=43
grid <- expand.grid(pH=pH.grid, LL=LL.grid, Temp=43)
table5 <- round(matrix(predict(mod3, newdata=grid, type="response"), nrow=5, ncol=5), 4)
colnames(table5) <- paste("pH", pH.grid, collape=" ")
rownames(table5) <- paste("LL", LL.grid, collape=" ")
#kable(table, caption="LL=0")

# PROBABILITY TABLE: Temp=50
grid <- expand.grid(pH=pH.grid, LL=LL.grid, Temp=50)
table6 <- round(matrix(predict(mod3, newdata=grid, type="response"), nrow=5, ncol=5), 4)
colnames(table6) <- paste("pH", pH.grid, collape=" ")
rownames(table6) <- paste("LL", LL.grid, collape=" ")
#kable(table, caption="LL=30")
kable(cbind(table5, table6), caption="Temp=43 and Temp=50")
```
